On the Functionality of the Multilayer Perceptron

### 2020-09-06 Thesis/Sections/FeatureSelections.tex
6 Sept 2020
- Updates to feature selections paper

@@

\paragraph*{}After frame-blocking, we apply a \textit{windowing function} to each frame. The analysis frame are concatenated into a $k \times N$ sized matrix, $A$, where each row is a frame. A standard \textit{Hann Window} of $N$ samples is generated and reshaped into a $N \times 1$ array,$H$. The $n$-th index in a Hann window with $N$ samples is defined:
\begin{equation}
\label{eqn-Hann}
H[n] = \frac{1}{2}\bigg[ 1 - \cos\Big( \frac{2\pi n}{N-1}\Big)\bigg]
\end{equation}
The window function is applied to each analysis frame by computing the matrix-product of the $k \times N$ block-frame matrix and the $N \times 1$ window function array. \textcolor{red}{Math is wrong here! Check this, check the code too!}

\paragraph*{}

@@

@@

\bibitem{Olsen}
Olson, Harry E. \textit{Music, Physics and Engineering}. 2nd ed., Dover Publications, 1967.

>>>>

\paragraph*{}Preprocessing a data set is a necessary step to execute prior to feature extraction \cite{Geron2,James}. In the case of audio files, preprocessing usually consists of ensuring that the data set contains the following:
\begin{enumerate}
\item A suitably sized number of files, of reasonable audio quality.
\item Audio encoded in a standard, and consistent format
\item A consistent sample rate between audio files
\item A consistent number of channels
\end{enumerate}
Note that different projects may require a different set of requirement from preprocessing\cite{Virtanen}. For this project, we have chosen to use the following parameters:
\begin{enumerate}
\item Roughtly $4000$ audio files Professionally or semi-professionally recorded in a studio. \textcolor{red}{Citation needed!}
\item All audio has be converted into \textit{.WAV} files from other formats, such as \textit{.AIF} or \textit{.MP3} using a MATLAB program
\item All audio is sampled at $44,100$ Hz
\item All audio has been down-mixed into mono-channel waveforms.
\end{enumerate}

@@

\paragraph*{}After frame-blocking, we apply a \textit{windowing function} to each frame. The analysis frame are concatenated into a $k \times N$ sized matrix, $A$, where each row is a frame. A standard \textit{Hann Window} of $N$ samples is generated and reshaped into a $N \times 1$ array,$H$. The $n$-th index in a Hann window with $N$ samples is defined:
\begin{equation}
\label{eqn-Hann}
H[n] = \frac{1}{2}\bigg[ 1 - \cos\Big( \frac{2\pi n}{N-1}\Big)\bigg]
\end{equation}
The window function is applied to each analysis frame by computing the dot-product of the Hann window array, $H$ and each row of the analysis frames matrix, $A$. This way, the window function has been effectively applied to each frame

\paragraph*{}Finally, we perform a \textit{Discrete Fourier Transform} (DFT) to bring the signal from a time domain into a frequency domain. \cite{Olson,Peatross}. The Discrete Fourier Transform is applied by producing an $N \times N$ \textit{transform matrix}, often noted at $\mathbb{W}$. Let $\omega^k = e^{\frac{-2\pi i}{N}k}$, then the DFT matrix for a time-space containing $N$ samples
\begin{equation}
\label{eqn-DFTMatrix}
\mathbb{W} = \frac{1}{\sqrt{N}}
\begin{bmatrix}
1 & 1 & 1 & 1 & \hdots & 1 \\
1 & \omega		& \omega^2 & \omega^3 & \hdots & \omega^{N-1} \\
1 & \omega^2	& \omega^4 & \omega^6 & \hdots & \omega^{2(N-1)} \\
1 & \omega^3	& \omega^6 & \omega^9 & \hdots & \omega^{3(N-1)} \\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots \\
1 & \omega^{N-1} & \omega^{2(N-1)} & \omega^{3(N-1)} & \hdots & \omega^{(N-1)^2}
\end{bmatrix}
\end{equation}
Each column of the matrix is a complex sinusoidal oscillating with an integer number of periods within the $N$-sample length window \cite{Short,Peatross}.The DFT is applied by a taking the matrix -product of $\mathbb{W}$ and $A^T$. The transpose of $A$ then makes each analysis frame into a column vector, which gives the appropriate dimension from multiplciation.

\paragraph*{}Most standard implementations of neural network require all activations to be real floating-point numbers. Since the the DFT matrix introduce complex values, we  compute the Hadamard (element-wise) product of $\mathbb{W}A^T$ and it's complex conjugate as such:
\begin{equation}
\label{eqn-Spectrogram}
\big( \mathbb{W}A^T \big) \odot \big( \mathbb{W}A^T \big)^* = S_{xx}
\end{equation}
Where $\mathbb{W}$ is the DFT matrix from eqn. (\ref{eqn-DFTMatrix}), $A^T$ is the transpose of the analysis frames matrix, and $^*$ denotes the element-wise complex conjugate. The matrix $S_{xx}$, is the spectrogram representation of the initial waveform, and now contains $N \times k$ real floating-point numbers.

\paragraph*{}Each Column of the $S_{xx}$ matrix is now a single-frame that has been moved into a frequency-space representation, thus there are $k$ columns, just as there were $k$ analysis frames. Given the discretized nature of digital audio, the fequency-sapce representation is not a continuous function, but rather a column vector, where the frequency has been assigned to one of $N$ bins. Since human hearing extends from about $20$ Hz to around $20$ kHz, audio files, such as music and voice recordings are typically sampled around $44.1$ kHz, to ensure that the full range of audio is held by the recording \cite{Olson,Virtanen}.

\paragraph*{}However, standard musical instruments seldom extend above $8$ kHz \cite{Olson,White}. This means that when constructing the spectrogram, we will rarely ever see energy present above this frequency, and the $S_{xx}$ will contain mostly zero, or zero-like entries. As a result of this, we can simply crop each matrix, to display frequencies that are only between $0$ Hz and $8000$ Hz. This makes the input smaller, and eliminates redundant and non-useful information. This brings the number rows in the $S_{xx}$ matrix from $N$ to $N'$. lastly, each raw audio file may contain a different number of samples to begin with, thus a different number of analysis frames are created. To ensure each sample presented to the classifier, we assert that each matrix must have exactly $k'$ columns. If $k < k'$, columns of zeros are added to pad the input, and if $k' < k$, columns are removed.

\paragraph*{}Each spectrogram is now $N' \times k'$ and effectively encodes the energy distribution of the waveform a a function of both time and frequency. The spectrogram is the first feature used in this model. For this classifier, we have chosen $N' = 560$ and $k' = 256$. For training, a mini-batch of $b$ samples are concatenated into a single array object. \textcolor{blue}{See Neural Network section for more details}. For a batch of $b$ samples of $N' \times k' \times 1$ spectrograms, we shape $X_1$ such that:
\begin{equation}
\label{eqn-X1 shape}
X_1 = \big\{ S_{xx}^{(0)},S_{xx}^{(1)},S_{xx}^{(2)}, ... , S_{xx}^{(b-1)} \big\} \in \mathbb{R}^{(b \times N' \times k' \times 1)}
\end{equation}

\paragraph*{}The fourth dimension, which contains $1$ element, allows for the spectrogram matrix to be processed identically to how neural networks process gray-scale images. For example, and RGB image would have a last-axis shape of $3$ elements. This matrix is presented to the \textit{Convolution} branch of the neural network from processing. \textcolor{blue}{See Neural Network section for more details}

@@

\paragraph*{}The features described in this section are derived from time-domain representations of each audio sample. In some cases, this can be from the waveform.

@@

\bibitem{Olson}
Olson, Harry E. \textit{Music, Physics and Engineering}. 2nd ed., Dover Publications, 1967.
